---
title: "Visual Analytics Project"
description: |
  I will explore the health of various employers in the city of Engagement, Ohio USA.
author:
  - name: Che Xuan 
    url: https://www.linkedin.com/in/jacob-che-xuan-b646a9123/
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.retina = 3)
```

# Overview

In this project, appropriate static and interactive statistical graphics methods are used to reveal the employment patterns of the [city of Engagement, Ohio USA](https://vast-challenge.github.io/2022/).

The data are processed by using appropriate **tidyverse** family of packages and the statistical graphics are prepared using **ggplot2** and its extensions.


# Installing & Launching R Packages

Before we get started, it is important for us to ensure that the required R packages have been installed. If yes, we will load the R packages. If they have yet to be installed, we will install the R packages and load them onto R environment.

The chunk code below will do the trick.

```{r}
packages = c('tidyverse', 'lubridate', 'treemapify', 'ggiraph', 'plotly', 'DT', 'patchwork', 'gganimate', 'treemap', 'd3treeR')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}

```


# Importing Data

```{r echo=FALSE}
participants <- read_csv("C:/Jacobche/ISSS608/th_ex/rawdata/Participants.csv")
checkin <- read_csv("C:/Jacobche/ISSS608/th_ex/rawdata/CheckinJournal.csv")
travel <- read_csv("C:/Jacobche/ISSS608/th_ex/rawdata/TravelJournal.csv")
employers <- read_csv("C:/Jacobche/ISSS608/th_ex/rawdata/Employers.csv")
jobs <- read_csv("C:/Jacobche/ISSS608/th_ex/rawdata/Jobs.csv")
```

# Data Wrangling

```{r echo=FALSE}
checkin_work <- checkin %>%
  filter(venueType == 'Workplace')
```

Percentage of jobs which involve weekends duties:

```{r echo=FALSE}
jobs_weekends <- jobs %>%
  mutate(day_type = ifelse(str_detect(daysToWork, c("Saturday", "Sunday")) == TRUE, "weekends", "weekdays")) %>%
  group_by(day_type) %>%
  summarise(count = n()) %>%
  ungroup()
```

```{r echo=FALSE}
ggplot(jobs_weekends, 
       aes(x = 2, y = count, fill = day_type)) +
  geom_col() +
  geom_text(aes(label=paste0(count, " (", round(count/sum(count)*100, 1), "%)")),
            position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y") +
  xlim(.1,2.5) +
  guides(fill = guide_legend(title = "Day Type")) +
  theme(panel.background = element_rect(fill = "white"),
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())
```

Look at by employers

```{r echo=FALSE}
jobs_employers <- jobs %>%
  group_by(employerId) %>%
  summarise(jobs_count = n(),
            avg_hourly_rate = round(mean(hourlyRate),1)) %>%
  ungroup()
```

```{r echo=FALSE}
jobs_employers$employerId = as.character(jobs_employers$employerId)
```


```{r echo=FALSE}
ggplot(jobs_employers, aes(area = avg_hourly_rate, fill = avg_hourly_rate)) +
  geom_treemap() +
  scale_fill_distiller(direction = 1)
```

By Company Size

```{r echo=FALSE}
jobs_employers_1 <- jobs_employers %>%
  group_by(jobs_count) %>%
  summarise(companies_count = n()) %>%
  ungroup() %>%
  mutate(percentage = round(companies_count/sum(companies_count)*100, 1))
```

```{r echo=FALSE}
jobs_employers_1$jobs_count = as.character(jobs_employers_1$jobs_count)
```

```{r echo=FALSE}
ggplot(jobs_employers_1, 
       aes(x = reorder(jobs_count, companies_count), y=companies_count)) +
  geom_bar(stat = "identity", color="black", fill="light blue") +
  geom_text(aes(label=paste0(companies_count, " (", round(companies_count/sum(companies_count)*100, 1), "%)")),
            position = position_stack(vjust = 0.5)) +
  coord_flip() +
  labs(y= 'Number of Companies', x= 'Jobs Count') +
  theme(axis.title.y= element_text(angle=0), axis.ticks.x= element_blank(), axis.text.x=element_blank(),
      axis.ticks.y= element_blank(), axis.line= element_line(color= 'grey'))
```


```{r}
write.csv(jobs_employers,"jobs_employers.csv", row.names = FALSE)
```

```{r}
write.csv(jobs_employers_1,"jobs_employers_1.csv", row.names = FALSE)
```


Travel to work

```{r}
travel$year_month <- format(as.Date(travel$`travelEndTime`), "%Y-%m")
travel$day <- day(travel$`travelEndTime`)
travel$wkday <- weekdays(travel$`travelEndTime`)
```

```{r}
travel_to_work <- travel %>%
  filter(purpose == 'Work/Home Commute') %>%
  inner_join(y=employers, by = c("travelEndLocationId" = "employerId")) %>%
  select(participantId, travelEndTime, year_month, day, wkday, travelEndLocationId, purpose, location, buildingId) %>%
  rename('employerId' = 'travelEndLocationId')
```

```{r}
travel_to_work$purpose <- sub('Work/Home Commute', 
                              'Work',
                              travel_to_work$purpose)
```

```{r}
travel_to_work_daily_count <- travel_to_work %>%
  group_by(employerId, year_month, day, wkday) %>%
  summarise(count = n()) %>%
  ungroup()
```

```{r}
travel_to_work_monthly_change <- travel_to_work_daily_count %>%
  group_by(employerId, year_month) %>%
  summarise(monthly_employees = max(count)) %>%
  mutate(mom_change = coalesce(monthly_employees - lead(monthly_employees),0),
         mom_turnover_rate = coalesce((monthly_employees - lead(monthly_employees))/monthly_employees,0)) %>%
  ungroup()
```

```{r}
turnover_by_employers <- travel_to_work_monthly_change %>%
  group_by(employerId) %>%
  summarise(turnovers = sum(mom_change)) %>%
  ungroup()
```

```{r}
employers_revised <- employers %>%
  inner_join(y=turnover_by_employers, by = c("employerId" = "employerId"))
```

```{r}
employers_revised$level <- cut(employers_revised$turnovers, breaks = c(-1,0,2,15), labels = c("Low Rate","Medium Rate","High Rate"))
```


```{r}
write.csv(employers_revised,"employers_revised.csv", row.names = FALSE)
```

















